---
title: "Class 19 HW"
author: "Matt Hashimoto"
date: "12/6/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Population Scale Analysis

#### Q13

"Read this file into R and determine the sample size for each genotype and their 
corresponding median expression levels for each of these genotypes."

```{r}
# Read file from url
finaldata <- read.table(url("https://bioboot.github.io/bggn213_W19/class-material/rs8067378_ENSG00000172057.6.txt"))

# Check summary of the overall data
summary(finaldata)
```

```{r}
# Create new tables based on genotype
AAtable <- finaldata[-which(finaldata$geno != "A/A"),]
AGtable <- finaldata[-which(finaldata$geno != "A/G"),]
GGtable <- finaldata[-which(finaldata$geno != "G/G"),]

# Check new tables
summary(AAtable)
summary(AGtable)
summary(GGtable)
```

The sample size of A/A is 108, A/G is 233, and G/G is 121. The median expression
level for A/A is 31.25, A/G is 25.065, G/G is 20.074.

#### Q14

"Generate a boxplot with a box per genotype, what could you infer from the
relative expression value between A/A and G/G displayed in this plot? Does the
SNP effect the expression of ORMDL3?"

```{r}
boxlist <- list("A/A" = AAtable$exp, "A/G" = AGtable$exp, "G/G" = GGtable$exp)
boxplot(boxlist,
        main = "Affect of Genotype on Expression Level in ORMDL3",
        xlab = "Genotype",
        ylab = "Expression Level",
        notch = TRUE,
        col = c("red", "green", "blue"))
```

From this box plot, it is reasonable to infer that genotype does significantly
affect expression levels of the ORMDL3 gene. The median value for the G/G data
falls well outside the IQR of the A/A data, indicating signficance. Thus, there
is strong evidence that the associated SNP affects the expression of ORMDL3. 





































