---
title: "Vaccination Rate Mini Project"
author: "Matt Hashimoto"
date: "12/3/2021"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Getting Started

Let's first start by loading our data from the .csv file:

```{r}
# Import vaccination data
vax <- read.csv("covid19vaccinesbyzipcode_test.csv")
head(vax)
```

#### Q1

"What column details the total number of people fully vaccinated?"

Column 9, titled "persons_fully_vaccinated".

#### Q2

"What column details the Zip code tabulation area?"

Column 2, titled "zip_code_tabulation_area".

#### Q3

"What is the earliest date in this dataset?"

This can be found by looking at the first entry in the "as_of_date" column:

```{r}
# View the first entry in the as_of_date column
vax$as_of_date[1]
```

Thus, the earliest date is January 5th, 2021.

#### Q4

"What is the latest date in this dataset?"

Similarly to the last question, this can be found by looking at the last entry
in the "as_of_date" column:

```{r}
vax$as_of_date[length(vax$as_of_date)]
```

Thus, the latest date is November 30th, 2021.

Let's try calling the skim function to get a better idea of what's in the
dataset:

```{r}
# Call the skim function
skimr::skim(vax)
```

#### Q5

"How many numeric columns are in this dataset?"

As seen from the skim results, there are 9 numeric columns.

#### Q6

"Note that there are “missing values” in the dataset. How many NA values are
there in the persons_fully_vaccinated column?"

The "n_missing" column shows that there are 8472 NA values in the
"persons_fully_vaccinated" column.

#### Q7

"What percent of persons_fully_vaccinated values are missing (to 2 significant
figures)?"

```{r}
# 8472 missing values out of 84672
8472 / 84672
```

10% of the values are missing.

#### Q8

"[Optional]: Why might this data be missing?"

This data may be missing because there is no method of collecting data from
specific zip codes. As mentioned earlier in the lab document, certain
institutions or organizations may have no obligation or reason to report their
vaccination data, and certain zip codes may be entirely managed by these
institutions or organizations.

## Working With Dates

Let's use the lubridate library to help us deal with dates:

```{r}
library(lubridate)
```

Check today's date:

```{r}
today()
```

Let's convert our dates into a lubridate format to make analysis easier:

```{r}
# Speciffy that we are using the Year-mont-day format
vax$as_of_date <- ymd(vax$as_of_date)
```

Now we can use lubridate functions to check things like how many days have
passed since the first data was collected:

```{r}
# Check time since first measurement
today() - vax$as_of_date[1]
```

We can also calculate how much time the data spans:

```{r}
# Check time span
vax$as_of_date[nrow(vax)] - vax$as_of_date[1]
```

#### Q9

"How many days have passed since the last update of the dataset?"

```{r}
today() - vax$as_of_date[nrow(vax)]
```

3 days have passed since the last update.

#### Q10

"How many unique dates are in the dataset (i.e. how many different dates are
detailed)?"

```{r}
length(unique(vax$as_of_date))
```

There are 48 unique dates in the dataset.

## Working With ZIP Codes

Let's load in the zipcodeR library:

```{r}
# Load the zipcodeR library
library(zipcodeR)
```

Next let's find the centroid of the 92037 zip code area (UCSD):

```{r}
# Find centroid of the 92037 zip code
geocode_zip('92037')
```

We can also calculate the distance between any two zip codes in miles:

```{r}
# Distance in miles
zip_distance('92037','92109')
```

We can also pull census data about zip codes:

```{r}
# Pull census data
reverse_zipcode(c('92037', "92109") )
```

We can use this to pull census data for all the zip codes we may be interested
in:

```{r}
# Pull data for all ZIP codes in the dataset
#zipdata <- reverse_zipcode( vax$zip_code_tabulation_area )
```

## Focus on the San Diego Area

We can restrict ourselves to San Diego county using base R:

```{r}
# Subset to San Diego county only areas
sd <- vax[vax$county == "San Diego",]
```

Or we could use the dplyr library:

```{r}
# Load library
library(dplyr)

# Filter just results from SD
sd <- filter(vax, county == "San Diego")
nrow(sd)
```

The dplyr package is more convenient when trying to subset across multiple
criteria:

```{r}
# All SD counties with populations over 10000
sd.10 <- filter(vax, county == "San Diego" &
                age5_plus_population > 10000)
```

#### Q11

"How many distinct zip codes are listed for San Diego County?"

```{r}
# Check for uniqueness
length(unique(sd$zip_code_tabulation_area))
```

107 distinct zip codes are listed for SD county.

#### Q12

"What San Diego County Zip code area has the largest 12 + Population in this
dataset?"

```{r}
# Check for max population value
sd$zip_code_tabulation_area[which.max(sd$age12_plus_population)]
```

The 92154 area has the largest 12+ population.

```{r}
# All data for Nov 16
sd.nov16 <- filter(vax, county == "San Diego" &
                as_of_date == "2021-11-16")
```

#### Q13

"What is the overall average “Percent of Population Fully Vaccinated” value for
all San Diego “County” as of “2021-11-16”?"

```{r}
# Average percent of population fully vaccinated
mean(sd.nov16$percent_of_population_fully_vaccinated, na.rm = TRUE)
```

The average percent of population fully vaccinated is 67.22%.

#### Q14

"Using either ggplot or base R graphics make a summary figure that shows the
distribution of Percent of Population Fully Vaccinated values as of
“2021-11-16”?"

```{r}
# Plot distribution of percent fully vaccinated
hist(sd.nov16$percent_of_population_fully_vaccinated,
     main = "Histogram of Vaccination Rates Across San Diego County",
     xlab = "Percent of Population Fully Vaccinated on 2021-11-16",
     col = "gray")
```

## Focus on UCSD/La Jolla

Let's filter to the UCSD area zip code:

```{r}
# Filter to UCSD zip code and check 5+ population
ucsd <- filter(sd, zip_code_tabulation_area == "92037")
ucsd[1,]$age5_plus_population
```

#### Q15

"Using ggplot make a graph of the vaccination rate time course for the 92037 ZIP
code area:"

```{r}
# Load ggplot library
library(ggplot2)

# Use ggplot to create a graph
ggplot(ucsd) +
  aes(ucsd$as_of_date,
      ucsd$percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group = 1) +
  ylim(c(0,1)) +
  labs(x = "Date", y = "Percent Vaccinated") +
  ggtitle("Vaccination Rate for La Jolla, CA 92037")
```

## Comparing 92037 to Other Similarly Sized Areas

Let's filter our vaccination data once again to data at least as large as the
population in 92037:

```{r}
# Subset to all CA areas with a population as large as 92037
vax.36 <- filter(vax, age5_plus_population > 36144 &
                as_of_date == "2021-11-16")

head(vax.36)
```

#### Q16

"Calculate the mean “Percent of Population Fully Vaccinated” for ZIP code areas
with a population as large as 92037 (La Jolla) as_of_date “2021-11-16”. Add this
as a straight horizontal line to your plot from above with the geom_hline()
function?"

```{r}
# Calculate mean
mean(vax.36$percent_of_population_fully_vaccinated, na.rm = TRUE)
```

```{r}
# Add line to plot
ggplot(ucsd) +
  aes(ucsd$as_of_date,
      ucsd$percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group = 1) +
  ylim(c(0,1)) +
  labs(x = "Date", y = "Percent Vaccinated") +
  ggtitle("Vaccination Rate for La Jolla, CA 92037") +
  geom_hline(yintercept = mean(vax.36$percent_of_population_fully_vaccinated,
                      na.rm = TRUE),
             color = "red", linetype = "dashed")
```

#### Q17

"What is the 6 number summary (Min, 1st Qu., Median, Mean, 3rd Qu., and Max) of
the “Percent of Population Fully Vaccinated” values for ZIP code areas with a
population as large as 92037 (La Jolla) as_of_date “2021-11-16”?"

```{r}
# Use fivenum to get min, 1st qu, median, 3rd qu, and max
fivenum(vax.36$percent_of_population_fully_vaccinated)

# Use mean()
mean(vax.36$percent_of_population_fully_vaccinated)
```

#### Q18

"Using ggplot generate a histogram of this data."

```{r}
ggplot(vax.36) +
  aes(vax.36$percent_of_population_fully_vaccinated) +
  geom_histogram() +
  xlim(0,1) +
  labs(x = "Percent Vaccinated")
```

#### Q19

"Is the 92109 and 92040 ZIP code areas above or below the average value you
calculated for all these above?"

```{r}
# The average value
mean(vax.36$percent_of_population_fully_vaccinated)

# Check 92109
vax %>% filter(as_of_date == "2021-11-16") %>%  
  filter(zip_code_tabulation_area=="92109") %>%
  select(percent_of_population_fully_vaccinated)

# Check 92040
vax %>% filter(as_of_date == "2021-11-16") %>%  
  filter(zip_code_tabulation_area=="92040") %>%
  select(percent_of_population_fully_vaccinated)
```

As you can see, the 92109 zip code is above the average vaccination percentage,
while the 92040 zip code is below.

#### Q20

"Finally make a time course plot of vaccination progress for all areas in the
full dataset with a age5_plus_population > 36144."

```{r}
# Filter data for all days
vax.36.all <- filter(vax, age5_plus_population > 36144)

# Plot with ggplot
ggplot(vax.36.all) +
  aes(vax.36.all$as_of_date,
      vax.36.all$percent_of_population_fully_vaccinated, 
      group = zip_code_tabulation_area) +
  geom_line(alpha = 0.2, color = "blue") +
  ylim(0,1) +
  labs(x = "Date", y = "Percent Vaccinated",
       title = "Vaccination Rate Across California",
       subtitle = "Only areas with a population above 36k are shown.") +
  geom_hline(yintercept = mean(vax.36$percent_of_population_fully_vaccinated,
                               na.rm = TRUE),
             linetype = "dashed")
```























